\section{Overview}

\begin{frame}{Data Quality: Why Preprocess the Data?}
	\begin{itemize}
		\item \textbf{Measures for {\color{airforceblue}data quality}: A
			      multidimensional view:}
		      \begin{itemize}
			      \item \textbf{Accuracy:} correct or wrong, accurate or not.
			      \item \textbf{Completeness:} not recorded, unavailable.
			      \item \textbf{Consistency:} some modified but some not, dangling
			            refs, etc.
			      \item \textbf{Timeliness:} timely updated?
			      \item \textbf{Believability:} how trustworthy is it, that the data
			            is correct?
			      \item \textbf{Interpretability:} how easily can the data be
			            understood?
			      \item And even many more!
		      \end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Major Tasks in Data Preprocessing (I)}
	\begin{itemize}
		\item \textbf{Data cleaning:}
		      \begin{itemize}
			      \item Fill in missing values.
			      \item Smooth noisy data.
			      \item Identify or remove outliers.
			      \item Resolve inconsistencies.
		      \end{itemize}
		\item \textbf{Data integration:}
		      \begin{itemize}
			      \item Integration of multiple databases.
			      \item Data cubes or files.
		      \end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Major Tasks in Data Preprocessing (II)}
	\begin{itemize}
		\item \textbf{Data reduction:}
		      \begin{itemize}
			      \item Dimensionality reduction.
			      \item Numerosity reduction.
			      \item Data compression.
		      \end{itemize}
		\item \textbf{Data transformation and data discretization:}
		      \begin{itemize}
			      \item Normalization.
			      \item Concept-hierarchy generation.
		      \end{itemize}
	\end{itemize}
\end{frame}
