\documentclass[
english,
smallborders
]{i6prcsht}
\usepackage{i6common}
\usepackage{i6lecture}

\usepackage{todonotes}
\usepackage[utf8]{inputenc}
\usepackage{textcomp}
\usepackage{pdfpages}
\usepackage{csquotes}
\usepackage{awesomebox}

\hyphenation{Stud-On}

\begin{document}

\title{Exercise Sheet 3: \\ Frequent Patterns}
\maketitle
\vspace*{-2cm}

\section*{About this Exercise Sheet}

This exercise sheet focuses on the content of the lecture \textit{6. Mining Frequent Patterns, Associations and Correlations}.

It includes both a practical data science exercise (Exercise 1) and theoretical exercises on Apriori (Exercise 2) and FP-Growth (Exercise 3).

The exercise sheet is designed for a two-week period, during which the tasks can be completed flexibly (Exercise 1 is planned for the first exercise session, and Exercises 2 and 3 for the second session).

The sample solution will be published after the two weeks have elapsed.

\section*{Preparation}

Before participating in the exercise, you must prepare the following:

\begin{enumerate}
	\item \textbf{Install Python and pip on your computer}

	      \begin{itemize}
		      \item Detailed instructions can be found in \texttt{1-Introduction-Python-Pandas.pdf}.
	      \end{itemize}

	\item \textbf{Download provided additional files}

	      \begin{itemize}
		      \item Download \texttt{Additional-Files.zip} from StudOn
		      \item Extract it to a folder of your choice.
	      \end{itemize}

	\item \textbf{Install required Python packages}

	      \begin{itemize}
		      \item Open a terminal and navigate to the folder where you extracted the files.
		      \item Run the command \texttt{pip install -r requirements.txt} within the extracted additional files folder to install the required Python packages.
	      \end{itemize}


\end{enumerate}

\section*{Exercise 1: Mining Frequent Patterns}

This exercise comprises practical data science tasks and thus utilizes a Jupyter Notebook:

\begin{enumerate}
	\item Open \texttt{Mining-Frequent-Patterns.ipynb}.
	\item Take a look at the tasks (blue boxes) in the notebook and try to solve them.
\end{enumerate}

If you are unfamiliar with how to open a Jupyter Notebook, please refer to Exercise 1 of \texttt{1-Introduction-Python-Pandas.pdf}.

\begin{solution}
	The solution to the exercise can be found in the \texttt{Additional-Files.zip} of the Solution.
\end{solution}

\section*{Exercise 2: Apriori}

TODO

\section*{Exercise 3: FP-Growth}

TODO


\end{document}
