\section{Data Objects and Attribute Types}

\begin{frame}{Types of Data Sets}
	\begin{columns}
		\begin{column}{0.45\textwidth}
			\textbf{Records:}
			\begin{itemize}[noitemsep]
				\item Relational records.
				\item Data matrix, e.g. numerical matrix, crosstabs.
				\item Document data: text documents, \textbf{term-frequency vectors}. \tikzmark{n1}
				\item \textbf{Transaction data}. \tikzmark{n2}
			\end{itemize}
			\textbf{Graph and network:}
			\begin{itemize}[noitemsep]
				\item World wide web.
				\item Social of information networks.
				\item Molecular structures.
			\end{itemize}
		\end{column}

		\begin{column}{0.45\textwidth}  %%<--- here
			\begin{table}
				\begin{tabular}{|c|c|c|c|c|c|c|}
					\multicolumn{1}{c|}{}   & \rotatebox[origin=c]{270}{team} & \rotatebox[origin=c]{270}{couch} & \rotatebox[origin=c]{270}{play} & \rotatebox[origin=c]{270}{ball} & \rotatebox[origin=c]{270}{score} & \rotatebox[origin=c]{270}{game} \\ \hline
					\tikzmark{t1} Document1 & 3                               & 0                                & 5                               & 0                               & 2                                & 6                               \\ \hline
					Document2               & 0                               & 7                                & 0                               & 2                               & 1                                & 0                               \\ \hline
					Document3               & 0                               & 1                                & 0                               & 0                               & 1                                & 2                               \\
					\hline
				\end{tabular}\\[0.5cm]
				\begin{tabular} { | c | l |}
					\hline
					\textbf{TID}    & \textbf{Items}             \\
					\hline
					\tikzmark{t2} 1 & Bread, Coke, Milk          \\
					2               & Beer, Bread                \\
					3               & Beer, Coke, Diapers, Milk  \\
					4               & Beer, Bread, Diapers, Milk \\
					5               & Coke, Diapers, Milk        \\
					\hline
				\end{tabular}
			\end{table}
			\begin{tikzpicture}[remember picture,overlay]
				\path[draw=blue,thick,->]<1-> ([yshift=1mm]n1) -- (t1);
				\path[draw=blue,thick,->]<1-> ([yshift=1mm]n2) -- (t2);
			\end{tikzpicture}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Types of Data Sets}
	\textbf{Ordered data:}
	\begin{itemize}[noitemsep]
		\item Video data: sequences of images.
		\item Temporal data: time series.
		\item Sequential data: transaction sequences.
		\item Genetic sequence data.
	\end{itemize}
	\textbf{Spatial, image and multimedia:}
	\begin{itemize}[noitemsep]
		\item Spatial data: maps.
		\item Image data.
		\item Video data.
	\end{itemize}
\end{frame}

\begin{frame}{Important Characteristics of Structured Data}
	\textbf{Dimensionality}:\\
	Curse of dimensionality (sparse high-dimensional data spaces).\\[0.2cm]
	% make the example with the volume of a cube

	\textbf{Sparsity}:\\
	Only presence counts.\\[0.2cm]

	\textbf{Resolution}:\\
	Patterns depend on the scale.\\[0.2cm]

	\textbf{Distribution}:\\
	Centrality and dispersion.
\end{frame}

\begin{frame}{Data Objects}
	\textbf{Data sets are made up of data objects}.\\
	\textbf{A data object represents an entity}.\\[0.2cm]

	Examples:
	\begin{itemize}
		\item Sales database: customers, store items, sales.
		\item Medical database: patients, treatments.
		\item University database: students, professors, courses.
	\end{itemize}

	They are also called:\\
	Sampels, examples, instances, data points, objects, tuples, \ldots\\[0.2cm]

	\textbf{Data objects are described by attributes}:
	\begin{itemize}
		\item Database rows $\rightarrow$ data objects.
		\item Columns $\rightarrow$ attributes.
	\end{itemize}
\end{frame}

\begin{frame}{Attributes}
	\textbf{Attribute}:\\
	Sometimes also in other context: field, dimension, feature, variable, \ldots\\[0.2cm]
	A data field encodes the property of an entity or feature of a data object.\\
	E.g. \texttt{customer\_ID}, \texttt{name}, \texttt{address}.\\[0.5cm]

	\textbf{Types}:
	\begin{itemize}
		\item Nominal.
		\item Binary.
		\item Ordinal.
		\item Numerical:
		      \begin{itemize}
			      \item Interval scaled.
			      \item Ratio scaled.
		      \end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Attribute Types}
	\begin{itemize}
		\item \textbf{Nominal}:
		      \begin{itemize}
			      \item Categories, states, or "names of things".\\
			            E.g. \texttt{hair\_color} $= \{\text{auburn, black, blond, brown, grey, red, white}\}$.\\
			            Other examples: \texttt{marital\_status}, \texttt{occupation}, \texttt{ID}, \texttt{ZIP code}.
		      \end{itemize}
		\item \textbf{Binary}:
		      \begin{itemize}
			      \item Nominal attribute with only two states ($0$ and $1$).
			      \item \textbf{Symmetric binaries}: both outcomes equally important, such as \texttt{gender}.
			      \item \textbf{Asymmetric binary}: outcomes not equally important. \\
			            E.g. medical test (positive vs. negative).\\
			            Convention: assign 1 to most important outcome (e.g. HIV positive).
		      \end{itemize}
		\item \textbf{Ordinal}:
		      \begin{itemize}
			      \item Values have a meaningful order (ranking),\\
			            but magnitude between successive values is not known.\\
			            E.g. \texttt{size} $= \{\text{small, medium, large}\}$, grades, army rankings.
		      \end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Numerical Attribute Types}
	\begin{itemize}
		\item \textbf{Numerical: Quantity (integer- or real-valued)}.
		\item \textbf{Interval scaled}:
		      \begin{itemize}
			      \item Measured on a scale of \textbf{equally sized} units.
			      \item Values have order.\\
			            E.g. temperature in $C$ or $F$, calender dates.
			      \item No true zero-point.
		      \end{itemize}
		\item \textbf{Ratio scaled}:
		      \begin{itemize}
			      \item Inherent \textbf{zero point}.
			      \item We can speak of values as being an order of magnitude larger \\
			            than the unit of measurement.\\
			            E.g. $10 K$ is twice as high as $5 K$.\\
			            E.g. temperature in Kelvin, length, counts, monetary quantities.
		      \end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Discrete vs. Continuous Attributes }
	\begin{itemize}
		\item \textbf{Discrete attribute}:
		      \begin{itemize}
			      \item Has finite or countably infinite elements.\\
			            E.g. ZIP code, profession, or the set of words in a collection of documents.
			      \item Sometimes represented as integer variables.
			      \item Note: Binary attributes are a special case of discrete attributes.
		      \end{itemize}
		\item \textbf{Continuous attribute}:
		      \begin{itemize}
			      \item Has real numbers as attribute values.\\
			            E.g. temperature, height, or weight.
			      \item Practically, real values can only be measured and represented using a finite number of digits.
			      \item Continuous attributes are typically represented as floating-point variables.
		      \end{itemize}
	\end{itemize}
\end{frame}
