\section*{Appendix}

\begin{frame}{Apriori Algorithm - Pseudo Code}
	$C_k$: candidate itemsets of size $k$\\
	$L_k$: frequent itemsets of size $k$\\[0.5cm]

	$L_1 = \{\text{frequent items}\}$;\\[0.5cm]
	\textbf{for} $(k=1; L_k \neq \emptyset; k\texttt{++})$ \textbf{do begin}\\
	\hspace{1cm} $C_{k+1} = \text{candidates generated from } L_k;$\\
	\hspace{1cm} \textbf{for each} transaction $t$ in database \textbf{do}\\
	\hspace{2cm} increment the count of all candidates in $C_{k+1}$ that are
	contained in $t$;\\[0.1cm]
	\hspace{1cm}$L_{k+1} = \text{candidates in } C_{k+1} \text{ with
			min\_sup};$\\
	\textbf{end;}\\[0.1cm]
	\textbf{return} $\bigcup_k L_k$;
\end{frame}

\begin{frame}{Other Improved Mining Methods}
	\begin{itemize}
		\item \textbf{AFOPT} (Liu et al., KDD'03)
		      \begin{itemize}
			      \item A "push-right" method for mining condensed frequent-pattern
			            (CFP) tree.
		      \end{itemize}
		\item \textbf{Carpenter} (Pan et al., KDD'03)
		      \begin{itemize}
			      \item Mine datasets with small rows but numerous columns.
			      \item Construct a row-enumeration tree for efficient mining.
		      \end{itemize}
		\item \textbf{FP-growth+} (Grahne \& Zhu, FIMI'03)
		      \begin{itemize}
			      \item Efficiently using prefix-trees in mining frequent itemsets.
		      \end{itemize}
		\item \textbf{TD-Close} (Liu et al., SDM'06)
	\end{itemize}
\end{frame}

\begin{frame}{Extension of Pattern-growth Mining Methodology}
	\begin{itemize}
		\item \textbf{Mining closed frequent itemsets and max-patterns.}
		      \begin{itemize}
			      \item CLOSET (DMKD'00), FPclose, and FPMax (Grahne \& Zhu, FIMI'03)
		      \end{itemize}
		\item \textbf{Mining sequential patterns.}
		      \begin{itemize}
			      \item PrefixSpan (ICDE'01), CloSpan (SDM'03), BIDE (ICDE'04)
		      \end{itemize}
		\item \textbf{Mining graph patterns.}
		      \begin{itemize}
			      \item gSpan (ICDM'02), CloseGraph (KDD'03)
		      \end{itemize}
		\item \textbf{Constraint-based mining of frequent patterns.}
		      \begin{itemize}
			      \item Convertible constraints (ICDE'01), gPrune (PAKDD'03)
		      \end{itemize}
		\item \textbf{Computing iceberg data cubes with complex measures.}
		      \begin{itemize}
			      \item H-tree, H-cubing, and Star-cubing (SIGMOD'01, VLDB'03)
		      \end{itemize}
		\item \textbf{Pattern-growth-based clustering.}
		      \begin{itemize}
			      \item MaPle (Pei et al., ICDM'03)
		      \end{itemize}
		\item \textbf{Pattern-growth-based classification.}
		      \begin{itemize}
			      \item Mining frequent and discriminative patterns (Cheng et al.,
			            ICDE'07)
		      \end{itemize}
	\end{itemize}
\end{frame}
