\section{What kind of applications are targeted?}

\begin{frame}{Applications of Data Mining (I)}
	\begin{itemize}
		\item \textbf{Wherever there is data and more knowledge is desired, 
		there are data mining applications.}\\
		\item Typical data mining applications:
		\begin{itemize}
			\item \textbf{Business Intelligence} \\
				  \small{Provides historical, current, and predictive views of 
				  business operation.}
			\item \textbf{Web Search Engines} \\
				  \small{Need to decide which pages to index, which ones to 
				  index and how to rank them for search.}
			\item \textbf{Fraud detection} \\
				  \small{Possible fraud attempts automatically based on 
				  suspicious patterns in transactions.}
			\item \textbf{Predictive Maintenance} \\
				  \small{Evaluation of sensor data to maintain machines in time 
				  before a defect occurs.}
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Applications of Data Mining (II)}
	\begin{itemize}
		\item Example research projects using data mining at FAU\footnote{Found 
		in the FAU CRIS (Current research information system): 
		\url{https://cris.fau.de/}}:
		\begin{itemize}
			\item \textbf{Prediction of product properties using data mining 
			methods} \\
				  \small{Prof. Dr.-Ing. Sandro Wartzack (Chair of Engineering 
				  Design)}
		 	\item \textbf{Combustion and fuel optimization for the utilization 
		 	of residues in biomass furnaces} \\
		 		  \small{Prof. Dr.-Ing. Jürgen Karl (Chair of Energy Process 
		 		  Engineering)}
		 	\item \textbf{CoralTrace – A new approach to understanding 
		 	climate-induced reef crises} \\
		 	      \small{Prof. Dr. Wolfgang Kießling (Chair of Palaeontology)}
		 	\item \textbf{Performance Analysis in Team Sports} \\
		 		  \small{Prof. Dr. Björn Eskofier (Machine Learning and Data 
		 		  Analytics Lab)}
	 		\item \textbf{And many more ...} \\
	 			  \small{The chair of computer science 6 (data management) 
	 			  has some projects related to data mining, too. More 
	 			  information will be given in the last lecture.}
		\end{itemize}
	\end{itemize}
\end{frame}