\section*{Appendix}

\begin{frame}{Evolution of Sciences (I)}
	\begin{itemize}
		\item Before $1600$, era of \textbf{empirical science}.
		\item $1600-1950$s, rise of \textbf{theoretical science}.
		      \begin{itemize}
			      \item Each discipline has grown a theoretical component.
			      \item Theoretical models often motivate experiments and generalize
			            our understanding.
		      \end{itemize}
		\item $1950-1990$s, rise of \textbf{computational science}.
		      \begin{itemize}
			      \item Over the last $50$ years most disciplines have grown a third,
			            computational branch.
			            \begin{itemize}
				            \item E.g. empirical, theoretical, and computational ecology.
				            \item E.g. physics, linguistics or biology.
			            \end{itemize}
			      \item Computational science traditionally meant simulation.
			      \item It grew out of our inability to describe reality by
			            closed-form mathematical models.
		      \end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Evolution of Sciences (II)}
	\begin{itemize}
		\item $1990-$now, rise of \textbf{data science}.
		      \begin{itemize}
			      \item The flood of data from new instruments and modern simulations.
			      \item The ability to economically store and manage petabytes of
			            data.
			      \item The internet makes all these archives world wide accessible.
			      \item Scientific \emph{information management}, \\
			            acquisition,\\
			            organization, \\
			            query, and \\
			            visualization scale almost linearly with amount of data.
			      \item \textbf{Data mining} is a major new challenge!
		      \end{itemize}
		\item For further reading:\\
		      \small{Jim Gray and Alex Szaly: \emph{The World Wide Telescope: An
				      Archetype for Online Science}, \\ Communications of the ACM
			      45(11):
			      50-54, 2002.}
	\end{itemize}
\end{frame}

\begin{frame}{Evolution of Sciences (III)}
	\begin{itemize}
		\item $1960$s: Data collection, database creation, \\
		      \hspace{1cm} integrated management systems (IMS), and \\
		      \hspace{1cm} network database management systems (DBMS).
		\item $1970$s: Relational data model, relational DBMS implementation
		      (RDBMS).
		\item $1980$s: RDBMS products, database creation, \\
		      \hspace{1cm} advanced data models (extended relational, object
		      oriented, deductive etc.),\\
		      \hspace{1cm} application-oriented DBMS (spatial, scientific,
		      engineering etc.).
		\item $1990$s: Data mining, data warehousing, multimedia databases, web
		      databases.
		\item $2000$s: Stream data management and mining,\\
		      \hspace{1cm} data mining and applications, \\
		      \hspace{1cm} web technology (XML, data integration), and global
		      information systems.
	\end{itemize}
\end{frame}

\begin{frame}{A Brief History of Data Mining Society}
	\begin{itemize}
		\item \textbf{1989 IJCAI Workshop on Knowledge Discovery in
			      Databases:}\\
		      Knowledge Discovery in Databases (G. Piatetsky-Shapiro and W. Frawley,
		      1991).
		\item \textbf{1991-1994 Workshops on Knowledge Discovery in
			      Databases:}\\
		      Advances in Knowledge Discovery and Data Mining (U. Fayyad, G.
		      Piatetsky-Shapiro, P. Smyth and R. Uthurusamy, 1996).
		\item \textbf{1995-1998 International Conferences on Knowledge
			      Discovery in Databases and Data Mining (KDD’95-98):}\\
		      Journal of Data Mining and Knowledge Discovery (1997).
		\item \textbf{ACM SIGKDD conferences since 1998 and SIGKDD
			      Explorations.}\\
		\item \textbf{More conferences on data mining:}\\
		      PAKDD (1997), PKDD (1997), SIAM-Data Mining (2001), (IEEE) ICDM (2001),
		      etc.
		\item \textbf{Journal ACM Transactions on KDD starting in 2007}.
	\end{itemize}
\end{frame}

\begin{frame}{Conferences and Journals on Data Mining (I)}
	\textbf{KDD Conferences:}
	\begin{itemize}
		\item ACM SIGKDD Int. Conf. on Knowledge Discovery in Databases and
		      Data Mining (KDD).
		\item SIAM Data Mining Conf. (SDM).
		\item (IEEE) Int. Conf. on Data Mining (ICDM).
		\item European Conf. on Machine Learning and Principles and Practices
		      of Knowledge Discovery and Data Mining (ECML-PKDD).
		\item Pacific-Asia Conf. on Knowledge Discovery and Data Mining (PAKDD).
		\item Int. Conf. on Web Search and Data Mining (WSDM).
	\end{itemize}
\end{frame}

\begin{frame}{Conferences and Journals on Data Mining (II)}
	\textbf{Other related conferences:}
	\begin{itemize}
		\item DB conferences: ACM SIGMOD, VLDB, ICDE, EDBT, ICDT, \ldots
		\item Web and IR conferences: WWW, SIGIR, WSDM, \ldots
		\item ML conferences: ICML, NIPS, ICLR \ldots
		\item PR conferences: CVPR, ICPR \ldots
	\end{itemize}
	\textbf{Journals:}
	\begin{itemize}
		\item Data Mining and Knowledge Discovery (DAMI or DMKD).
		\item IEEE Trans. On Knowledge and Data Eng. (TKDE).
		\item KDD Explorations.
		\item ACM Trans. on KDD.
	\end{itemize}
\end{frame}

\begin{frame}{Starting points to Find References? (I)}
	\textbf{Data mining and KDD (SIGKDD: CD-ROM):}
	\begin{itemize}
		\item Conferences: ACM-SIGKDD, IEEE-ICDM, SIAM-DM, PKDD, PAKDD, etc.
		\item Journal: Data Mining and Knowledge Discovery, KDD Explorations,
		      ACM TKDD.
		\item KDnuggets: \url{www.kdnuggets.com}.
	\end{itemize}
	\textbf{Database systems (SIGMOD: ACM SIGMOD Anthology CD-ROM):}
	\begin{itemize}
		\item Conferences: ACM-SIGMOD, ACM-PODS, VLDB, IEEE-ICDE, EDBT, ICDT,
		      DASFAA.
		\item Journals: IEEE-TKDE, ACM-TODS/TOIS, JIIS, J. ACM, VLDB J., Info.
		      Sys., etc.
	\end{itemize}
	\textbf{AI \& Machine Learning:}
	\begin{itemize}
		\item Conferences: Machine learning (ML), AAAI, IJCAI, COLT (Learning
		      Theory), CVPR, NIPS, etc.
		\item Journals: Machine Learning, Artificial Intelligence, Knowledge
		      and Information Systems, IEEE-PAMI, etc.
	\end{itemize}
\end{frame}

\begin{frame}{Starting points to Find References? (II)}
	\textbf{Web and IR:}
	\begin{itemize}
		\item Conferences: SIGIR, WWW, CIKM, etc.
		\item Journals: WWW: Internet and Web Information Systems.
	\end{itemize}
	\textbf{Statistics:}
	\begin{itemize}
		\item Conferences: Joint Stat. Meeting, etc.
		\item Journals: Annals of Statistics, etc.
	\end{itemize}
	\textbf{Visualization:}
	\begin{itemize}
		\item Conferences: CHI, ACM-SIGGraph, etc.
		\item Journals: IEEE Trans. Visualization and Computer Graphics, etc.
	\end{itemize}
\end{frame}
